A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-03-24 demystify.tex
updates

While you might not have control over LORs (\autoref{sec:lor}) or where your go to school (\autoref{sec:your-school}), you do over your
statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}! So write it well because it could make a big difference.
In your statement, you have the opportunity to make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
% is important because if you need (GTA) funding, it will provide evidence
% that you can teach and communicate with students.

@@

Finally, you should get your SOP reviewed by LoR writers, professors (especially those who have served in adcom), or even graduate PhD students (because they have been through this process).

\begin{commentbox}[Vu:]
  I often read the research statement and LORs first. If I am
  persuaded by then, I would skim over other factors and advocate for
  admission (unless I see red flags in other parts). However, if I am not
  convinced, then I will likely recommend rejection (unless I see
  something standout in other parts).
  \tcblower
  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty (see more about various types of faculty in \autoref{sec:faculty-types}).
  Also, be careful not to send statements to wrong schools or mixing
  facts (e.g., talking about school X but mentioned about working with
  profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
  should.
\end{commentbox}

@@

\paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
take standardized English test.

\textbf{Advice}: Just do well enough to pass minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
Just as with grades (\autoref{sec:your-school}) and GRE, having a high scores in these tests might not help, but having too low scores can a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.

@@

However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).  In this case, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.

@@

Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.  This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find adviser.

@@

Another effective way is taking graduate level courses in the topics you are interested in.  Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).

>>>>

While you might not have control over LORs (\autoref{sec:lor}) or where your go to school (\autoref{sec:your-school}), you do over your
statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}! So write it well because it could make a big difference.

In your statement, you have the opportunity to connect directly with the adcom and make your application stand out and unique. Your SOP can convince the adcom that you would \emph{fit} the program, even if you don't have very strong research experience.
A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
% is important because if you need (GTA) funding, it will provide evidence
% that you can teach and communicate with students.

@@

Finally, you should get your SOP reviewed by your LoR writers, professors (especially those who have served in adcom), or even graduate PhD students (because they have been through this process).

\begin{commentbox}[Vu:]
  I often read SOP and LORs first. If I am
  persuaded by then, I would skim over other factors and advocate for
  admission (unless I see red flags in other parts). However, if I am not
  convinced, then I will likely recommend rejection (unless I see
  something really standout in other parts).
  \tcblower
  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty (see more about various types of faculty in \autoref{sec:faculty-types}).
  Also, be careful not to send statements to wrong schools or mixing
  facts (e.g., talking about school X but mentioned about working with
  profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
  should.
\end{commentbox}

@@

\paragraph{English Tests} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
take standardized English test.

\textbf{Advice}: Just do well enough to pass minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
Just as with grades (\autoref{sec:your-school}) and GRE, having a high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.

@@

However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they you to do research for them, then they likely will interview you first).  In the case of no interviews, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.

@@

Fortunately, while some non-US programs require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.  This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find adviser.

@@

Another effective way is taking graduate level courses in the topics you are interested in.  Professors teaching \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).

### 2024-03-24 faculty.tex
updates

    % Year
    \draw[ultra thick,gray]
    ($(current page.center)+(5,2)$) -- ++(0,-3cm) 
    node[
    midway,
    left=0.25cm,
    text width=5cm,
    align=right,
    black!75
    ]
    {
    {\fontsize{25}{30} \selectfont \bf Computer\\[8pt] Science}
    } 
    node[
    midway,
    right=0.25cm,
    text width=6cm,
    align=left,
    ForestGreen]
    {
    {\fontsize{48}{50} \selectfont PhD}
    };

    % Title
    \node[align=center] at ($(current page.center)+(0,-5)$) 
    {
    {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
    {\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]

@@

You are expected to have certain types of services as a faculty member.  While some services are required (e.g., when your department chair assigns you to some committee), some are not and you can decline (e.g., an invitation to be a PC member).  You should be strategic about which services you accept.

\section{Service to your Professional Community}

These are services that you do for your professional community. They will help you to get to know people in your field, and to get your name out.  Common examples include serving on the program committees of conferences, serving on editorial boards of journal, reviewing papers for journals, serving on grant panels.

\subsection{Serving on Program Committees (PC)} You should try to serve on the PC  of the \emph{top} conferences in your field. This is a good way to get to know what is hot in your field, and to get to know people in your field.  It is also a good way to get your name out.

\textbf{Advice}: You can ask senior colleagues or even your PhD advisor/collaborators to recommend you to serve on a PC. You can also directly email the PC chairs and ask to serve on the PC.

\subsection{Serving on NSF Panels} You should serve on NSF panels.  It is a great way to learn about the grant process: how to distinguish between good and bad proposals and how people review proposals.
It is also a good way learn what other people are doing, what is hot, and what is not. In short, it is a good way to learn how to write successful proposals.

@@

\textbf{Advice}: Similarly to serving in PC, you can cold email program managers in your area and ask to be on a panel. Once you have served, you will likely be invited to serve again.  Unfortunately, I do not have much experience with non-NSF panels or proposals.

@@

\subsection{Personal Experience} I have served on many PCs and NSF panels. The initial services are mostly through ``cold'' emails. For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked). I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
I also emailed NSF program managers and asked to serve on a panel, and they accepted my request.

I felt that I didn't get much out of serving on PC and didn't really like it much. However, I still do it because I feel I should give back to the community.
On the otherhand, I throroughly enjoy serving on NSF panels. While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions. Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.

@@

\textbf{Advice}: You should try to be strategic about which committees you serve on that can help build your lab and tenure case. For example, serving in the PhD adcom is very common for new assistant professors, as it is a good way to recruit students for your lab. Serving on the faculty search committee is also a good way to recruit new faculty members that you like (and also let you know what the job market is like, just in canse you want to move). Serving on the colloquium committee is also a good way to invite speakers in your field that can be your \emph{tenure letter writers}.

\subsubsection{Personal} Since coming to GMU, I have been much more active in serving the department.  I share the vision of the department and want to help achieve it. In addition to the typical services load, I volunteer to serve in additional committees and leadership roles. For example, I maintain the CSRankings DB for the department, I help organize the Open House to recruit admitted students, I serve in the Dept Chair renewal committee, etc.
Even though I am not on the search committee, I speak to every candidate and help recruit strong ones. I agree to be director of the MS program, which is a lot of work and typically done by tenured faculty.

Sometimes I feel overwhelmed by the services, but I also feel that I am making a difference in the department and that is rewarding.
I am also building a good relationship with my colleagues and the department chair, and that is important for my tenure case.
In addition, I learn a lot from being in various roles (e.g., I learned how class scheduling work, how the deparmtnet is run, how to recruit students, how to recruit faculty, how to run a colloquium, etc) and feel more confident in my leadership skills.

In short, if you are enthusiastic about the department, want to make a difference, and \emph{feel comfortable at handling additional load}, you can volunteer for more services and leadership roles.

@@

>>>>

    % Year
    \draw[ultra thick,gray]
    ($(current page.center)+(5,2)$) -- ++(0,-3cm) 
    node[
    midway,
    left=0.25cm,
    text width=5cm,
    align=right,
    black!75
    ]
    {
    {\fontsize{25}{30} \selectfont \bf Computer\\[8pt] Science}
    } 
    node[
    midway,
    right=0.25cm,
    text width=6cm,
    align=left,
    ForestGreen]
    {
    {\fontsize{40}{42} \selectfont Faculty}
    };

    % Title
    \node[align=center] at ($(current page.center)+(0,-5)$) 
    {
    {\fontsize{24}{24} \selectfont {{Demystifying Tenure Track:}}} \\[0.15in]
    {\fontsize{18}{18} \selectfont {{Navigating the Faculty Journey In Computer Science}}} \\[1in]    
    %{\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]

@@

While only 10\% of your work is officially for services, they are very important for your tenure and career.
%You are expected to have certain types of services as a faculty member.  
Some services are required (e.g., when your department chair assigns you to some committee), but some are not and you can decline (e.g., an invitation to be a PC member).  You should be strategic about which services you accept.

\section{Service to your Research Community}

These are services that you do for your research community. They will help you to get to know people in your field, and to get your name out.  Common examples include serving on the program committees of conferences, serving on editorial boards of journal, and serving on grant panels.

\subsection{Serving on Program Committees (PCs) and Editorial Boards} You should try to serve on the PCs  of the \emph{top} conferences in your field. This is a good way to know how to write good papers, and get to know the literature, what is hot in your field, and the people (who do what). It is also a good way to get your name out.  Similarly to PCs, you should also try to serve on the editorial boards of top journals in your field.

Serving on PC (or editor boards) essentially means reviewing papers. You will have to read papers and write reviews, and then argue for why a paper should be accepted or not.  You will also have to discuss with other PC members (or editors) and come to a consensus.  Typically the review process can last for several weeks, and there are also rebuttals and conditionally acceptances (which you should already be familiar with as authors).

\subsection{Serving on NSF Panels} As mentioned, faculty in CS typically get most of their funding from NSF. Thus, you should serve on NSF panels to learn about the grant process: how to distinguish between good and bad proposals and how people review proposals.
It is also a good way learn what other people are doing, what is hot, and what is not. In short, it is a good way to learn how to write successful proposals.

@@

@@

\subsection{Advice} To serve, you can directly ask people to recommend you. For example, for PC, ask senior colleagues or even your PhD advisor/collaborators to recommend you to the PC chairs. You can also directly email the PC chairs and ask to serve.  Once you have served well, you will likely be invited to serve again.  
Similarly, for NSF Panels, you can cold email program managers in your area and ask to be on their panels.  Unfortunately, I do not have much experience with non-NSF panels or proposals. For leadership roles, you can directly ask the general chair of a conference to see if you can take more leadership roles.

\subsection{Personal Experience} I have served on many PCs and NSF panels. I mostly cold email people for these opportunities. For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked). I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
I also emailed NSF program managers and asked to serve on a panel, and they accepted my request.

I felt that I didn't get much out of serving on PC and didn't really like it much. However, I still do it because I feel I should give back to the community.
On the other hand, I thoroughly enjoy serving on NSF panels (though of course my proposals still got rejected frequently). While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions. Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.

@@

\subsection{Advice} You should try to be strategic about which committees you serve on that can help build your lab and tenure case. For example, serving in the PhD adcom is very common for new assistant professors, as it is a good way to recruit students for your lab. Serving on the faculty search committee is also a good way to recruit new faculty members that you like (and also let you know what the job market is like, what is hot and not). Serving on the colloquium committee is also a good way to invite speakers in your field that can be your \emph{tenure letter writers}.

\subsection{Personal} Since coming to GMU, I have been very active in serving the department.  I share the vision of the department and want to help achieve it. In addition to the typical services load, I volunteer to serve in additional committees and leadership roles. For example, I maintain the CSRankings DB for the department, organize the Open House to recruit admitted students, participate in the Dept Chair renewal committee, etc.
Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was my study break) and help recruit strong ones. I also agree to be director of the MS program, which is a lot of work and typically done by tenured faculty.

Sometimes I feel overwhelmed by the services, but I also feel that I am making a difference in the department and that is rewarding.
I am also building a good relationship with my colleagues (and also head butt with some), and that is important for my tenure case.
In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to run a colloquium, etc) and feel more confident in my leadership skills.

In short, if you want to make a difference, and \emph{feel comfortable} at handling additional load, you can volunteer for more services. Otherwise, you can just ``stay low'',
do the required services and focus on your research and teaching.

@@

\chapter{Tenure Package and Application}

\section{The Package}

\subsection{Statements}

\subsection{CV}

\subsection{Letters}

\subsection{Others}
